Equivalent Income Stuff - Energy Crisis Plots

### 2023-09-12 minos/outcomes/EI_PresPlots_SimpleBatch.Rmd
Changed loneliness to string var from numeric so its easier to understand. Also lots of changes to EI plots

```{r}
# S7.var.list <- c('hh_income', 'equivalent_income',  # Income variables
#                  'S7_housing_quality', 'S7_neighbourhood_safety', 'S7_physical_health', 'S7_mental_health', 'S7_labour_state', 'loneliness',  # SIPHER 7 variables
#                  'ethnicity', 'age', 'region', 'job_sec', 'education_state', 'nkids_ind', 'housing_tenure', 'urban')
# out.path.batch <- here::here('output', 'SIPHER7_batch/')
# 
# base.batch <- read_batch_out_all_years(out.path.batch, 'baseline', start.year = 2021, var.list = S7.var.list, verbose=FALSE)
# wage.batch <- read_batch_out_all_years(out.path.batch, 'livingWageIntervention', start.year = 2021, var.list = c('income_boosted', S7.var.list), verbose=FALSE)
# energy.batch <- read_batch_out_all_years(out.path.batch, 'energyDownlift', start.year = 2021, var.list = c('income_boosted', S7.var.list), verbose=FALSE)
# child.batch <- read_batch_out_all_years(out.path.batch, 'hhIncomeChildUplift', start.year = 2021, var.list = c('income_boosted', S7.var.list), verbose=FALSE)
# pov.batch <- read_batch_out_all_years(out.path.batch, 'hhIncomePovertyLineChildUplift', start.year = 2021, var.list = c('income_boosted', S7.var.list), verbose=FALSE)

out.path.arc <- here::here('output', 'SIPHER7_batch_ARC/')

#base.batch <- read.csv(paste0(out.path.arc, 'baseline_batch_aggregated.csv'))
#energy.batch <- read.csv(paste0(out.path.arc, 'energy_batch_aggregated.csv'))
```

```{r}
## second batch dump
out.path.arc2 <- here::here('output', 'batch')

#target.pattern <- paste0('aggregated_run_id_', year, '.csv')
base.list <- list.files(path = paste0(out.path.arc2, '/baseline'),
                            full.names = TRUE)
base.batch <- do.call(rbind, lapply(base.list, read.csv))

energy.list <- list.files(path = paste0(out.path.arc2, '/energy'),
                            full.names = TRUE)
energy.batch <- do.call(rbind, lapply(energy.list, read.csv))

#base.batch <- 
```

@@

base.batch <- calc_age_group(base.batch)
energy.batch <- calc_age_group(energy.batch)
```

@@

base.batch <- calc_income_quintile(base.batch)
energy.batch <- calc_income_quintile(energy.batch)
```

@@

```{r}
income_equivinc_change_batch(base.batch, energy.batch, 'Energy Crisis')
```

@@

```{r}
inc_change_batch_subpop(base.batch, energy.batch, 'Energy Crisis', group.var = 'income.quint')
equivinc_change_batch_subpop(base.batch, energy.batch, 'Energy Crisis', group.var = 'income.quint')
```

@@

  int.inc <- energy.batch %>%
    select(pidp, time, run_id, hh_income, equivalent_income, all_of(group.var), weight) %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    summarise(energy_income = weighted.mean(hh_income, w = weight, na.rm = TRUE),
              energy_equiv_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE))

@@

@@

income_equivinc_percent_change_batch(base.batch, int.batch, int.name, group.var = 'income.quint')
```

@@

```{r}
S7_ordinal_pathway_batch <- function(base.bat, int.bat, int.name, var) {
  base.hous <- base.bat %>%
    select(pidp, time, run_id, all_of(var), weight) %>%
    group_by(time, run_id, .data[[var]]) %>%
    count() %>%
    group_by(time, run_id) %>%
    mutate(total = sum(n)) %>%
    mutate(prop = n / total) %>%
    select(-n, -total) %>%
    mutate(source = 'Baseline')

  inc.hous <- int.bat %>%
    select(pidp, time, run_id, all_of(var), weight) %>%
    group_by(time, run_id, .data[[var]]) %>%
    count() %>%
    group_by(time, run_id) %>%
    mutate(total = sum(n)) %>%
    mutate(prop = n / total) %>%
    select(-n, -total) %>%
    mutate(source = int.name)

  combined <- rbind(base.hous, inc.hous)
  combined[[var]] <- as.factor(combined[[var]])

  #print(combined)

  p1 <- ggplot(combined, aes(x = time, y = prop, group = interaction(.data[[var]], source), color = .data[[var]], linetype = source)) +
    geom_smooth() +
    geom_point()

  print(p1)

  base.hous2 <- base.bat %>%
    select(pidp, time, run_id, all_of(var)) %>%
    group_by(time, run_id, .data[[var]]) %>%
    count() %>%
    group_by(time, run_id) %>%
    mutate(total = sum(n)) %>%
    mutate(prop = (n / total) * 100) %>%
    select(-n, -total) %>%
    rename(Baseline = prop)

  int.hous2 <- int.bat %>%
    select(pidp, time, run_id, all_of(var)) %>%
    group_by(time, run_id, .data[[var]]) %>%
    count() %>%
    group_by(time, run_id) %>%
    mutate(total = sum(n)) %>%
    mutate(prop = (n / total) * 100) %>%
    select(-n, -total) %>%
    rename(int = prop)

  combined2 <- merge(base.hous2, int.hous2, by = c('time', 'run_id', var))

  print(combined2)

  combined2[[var]] <- as.factor(combined2[[var]])

  combined2$diff <- combined2$int - combined2$Baseline

  p2 <- ggplot(combined2, aes(x = time, y = diff, group = .data[[var]], color = .data[[var]])) +
    geom_smooth() +
    geom_point()

  print(p2)

}

S7_ordinal_pathway_batch(base.batch, energy.batch, 'Energy Crisis', 'S7_housing_quality')
```

```{r}
## ATTEMPT 2

group.var <- 'income.quint'
group.level <- 1
var <- 'S7_housing_quality'

base.sub <- base.batch %>%
  select(pidp, time, run_id, equivalent_income, all_of(group.var), all_of(var), weight) %>%
  group_by(time, run_id, .data[[group.var]], .data[[var]]) %>%
  summarise(baseline_EI = weighted.mean(equivalent_income, w = weight))

int.sub <- energy.batch %>%
  select(pidp, time, run_id, equivalent_income, all_of(group.var), all_of(var), weight) %>%
  group_by(time, run_id, .data[[group.var]], .data[[var]]) %>%
  summarise(int_EI = weighted.mean(equivalent_income, w = weight))

combined <- merge(base.sub, int.sub, by = c('time', 'run_id', group.var, var))
combined$diff <- combined$int_EI - combined$baseline_EI

ggplot(data = combined, aes(x = time, y = diff, group = .data[[var]], color = .data[[var]])) +
  geom_smooth() +
  facet_wrap(~ .data[[group.var]])

base.sub2 <- base.batch %>%
  select(pidp, time, run_id, equivalent_income, all_of(group.var), all_of(var), weight) %>%
  filter(.data[[group.var]] == group.level) %>%
  group_by(time, run_id, .data[[var]]) %>%
  summarise(equivalent_income = weighted.mean(equivalent_income, w = weight))
```

@@

  int.sub <- energy.batch %>%
    select(pidp, time, run_id, equivalent_income, all_of(group.var), all_of(var)) %>%
    group_by(time, run_id, .data[[group.var]], .data[[var]]) %>%
    count() %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    mutate(total = sum(n),
           int_prop = n / total) %>%
    select(-n, -total)

@@

```{r}
S7_pathway_change_byGroup(base.batch, energy.batch, group.var = 'income.quint', var = 'S7_housing_quality')
S7_pathway_change_byGroup(base.batch, energy.batch, group.var = 'income.quint', var = 'S7_neighbourhood_safety')
S7_pathway_change_byGroup(base.batch, energy.batch, group.var = 'income.quint', var = 'S7_physical_health')
S7_pathway_change_byGroup(base.batch, energy.batch, group.var = 'income.quint', var = 'S7_mental_health')
S7_pathway_change_byGroup(base.batch, energy.batch, group.var = 'income.quint', var = 'loneliness')
```

@@

```{r}
S7_cumulative_pathwayChange_byGroup(base.batch, energy.batch, group.var = 'income.quint', var = 'S7_housing_quality')
S7_cumulative_pathwayChange_byGroup(base.batch, energy.batch, group.var = 'income.quint', var = 'S7_neighbourhood_safety')
S7_cumulative_pathwayChange_byGroup(base.batch, energy.batch, group.var = 'income.quint', var = 'S7_physical_health')
S7_cumulative_pathwayChange_byGroup(base.batch, energy.batch, group.var = 'income.quint', var = 'S7_mental_health')
S7_cumulative_pathwayChange_byGroup(base.batch, energy.batch, group.var = 'income.quint', var = 'loneliness')
```

@@

```{r}
S7_cumulative_pathwayChange_byGroup(base.batch, energy.batch, group.var = 'age.group', var = 'S7_housing_quality')
S7_cumulative_pathwayChange_byGroup(base.batch, energy.batch, group.var = 'age.group', var = 'S7_neighbourhood_safety')
S7_cumulative_pathwayChange_byGroup(base.batch, energy.batch, group.var = 'age.group', var = 'S7_physical_health')
S7_cumulative_pathwayChange_byGroup(base.batch, energy.batch, group.var = 'age.group', var = 'S7_mental_health')
S7_cumulative_pathwayChange_byGroup(base.batch, energy.batch, group.var = 'age.group', var = 'loneliness')
```

# SIMPLIFIED Positive v Negative Change

```{r}
group.var <- 'income.quint'
var <- 'S7_mental_health'

base.sub <- base.batch %>%
  select(pidp, time, run_id, all_of(group.var), all_of(var)) %>%
  rename(Baseline_var = S7_mental_health,
         Baseline_income.quint = income.quint)

int.sub <- energy.batch %>%
  select(pidp, time, run_id, all_of(group.var), all_of(var)) %>%
  rename(Intervention_var = S7_mental_health,
         Intervention_income.quint = income.quint)

#combined <- rbind(base.sub, int.sub)

```{r}
# move further analysis here so we don't have to do that expensive merge every time
merged2 <- merged %>%
  mutate(varChange = Intervention_var - Baseline_var,
         effect = case_when(varChange > 0 ~ "Positive",
                            varChange < 0 ~ "Negative",
                            varChange == 0 ~ "No_Change")) %>%
  select(-Baseline_var, -Intervention_var, -Intervention_income.quint, -varChange) %>%
  rename(income.quint = Baseline_income.quint) %>%
  group_by(time, run_id, .data[[group.var]], effect) %>%
  count() %>%
  group_by(time, run_id, .data[[group.var]]) %>%
  mutate(total = sum(n),
         effect_prop = n / total) %>%
  select(-n, -total) %>%
  group_by(time, .data[[group.var]], effect) %>%
  summarise(effect_prop = mean(effect_prop)) %>%
  pivot_wider(names_from = 'effect',
              values_from = 'effect_prop') %>%
  mutate(change = Positive - Negative) %>%
  select(-Positive, -Negative, -No_Change, -time) %>%
  group_by(.data[[group.var]]) %>%
  summarise(change = sum(change))

ggplot(data = merged2, aes(x = income.quint, y = change)) +
  geom_col()
```

@@

  if (var == 'S7_housing_quality') {
    base.batch$S7_housing_quality <- as.numeric(factor(base.batch$S7_housing_quality,
                                            levels = c('No to all',
                                                       'Yes to some',
                                                       'Yes to all'),
                                            labels = c(1, 2, 3)))
    int.batch$S7_housing_quality <- as.numeric(factor(int.batch$S7_housing_quality,
                                            levels = c('No to all',
                                                       'Yes to some',
                                                       'Yes to all'),
                                            labels = c(1, 2, 3)))
  } else if (var == 'S7_neighbourhood_safety') {
    base.batch$S7_neighbourhood_safety <- as.numeric(factor(base.batch$S7_neighbourhood_safety,
                                                 levels = c('Often',
                                                            'Some of the time',
                                                            'Hardly ever'),
                                                 labels = c(1, 2, 3)))
    int.batch$S7_neighbourhood_safety <- as.numeric(factor(int.batch$S7_neighbourhood_safety,
                                                 levels = c('Often',
                                                            'Some of the time',
                                                            'Hardly ever'),
                                                 labels = c(1, 2, 3)))
  } else if (var == 'loneliness') {
    base.batch$loneliness <- as.numeric(factor(base.batch$loneliness,
                                    levels = c(1, 2, 3),
                                    labels = c(3, 2, 1)))
    int.batch$loneliness <- as.numeric(factor(int.batch$loneliness,
                                    levels = c(1, 2, 3),
                                    labels = c(3, 2, 1)))
  }

@@

  p1 <- ggplot(data = merged2, aes(x = .data[[group.var]], y = sum.change)) +
    geom_boxplot(notch = FALSE) +
    labs(title = paste0('Change in ', var, ' by ', group.var)) +
    ylab('% Change')

@@

```{r}
simplified_component_change(base.batch, energy.batch, group.var = 'income.quint', var = 'S7_housing_quality', print = TRUE)
simplified_component_change(base.batch, energy.batch, group.var = 'income.quint', var = 'S7_neighbourhood_safety', print = TRUE)
simplified_component_change(base.batch, energy.batch, group.var = 'income.quint', var = 'S7_mental_health', print = TRUE)
simplified_component_change(base.batch, energy.batch, group.var = 'income.quint', var = 'S7_physical_health', print = TRUE)
simplified_component_change(base.batch, energy.batch, group.var = 'income.quint', var = 'loneliness', print = TRUE)
```

@@

df <- simplified_component_change(base.batch, energy.batch, group.var = 'income.quint', var = 'S7_housing_quality')
colnames(df) <- c('Income_Quintile', 'S7_housing_quality')
for (S7_var in varlist) {
  df2 <- simplified_component_change(base.batch, energy.batch, group.var = 'income.quint', var = S7_var)
  colnames(df2) <- c('Income_Quintile', S7_var)
  df <- merge(df, df2, by = c('Income_Quintile'))
}
rm(df2)

@@

df <- simplified_component_change_BOXPLOT(base.batch, energy.batch, group.var = 'income.quint', var = 'S7_housing_quality')
colnames(df) <- c('run_id', 'Income_Quintile', 'S7_housing_quality')
for (S7_var in varlist) {
  df2 <- simplified_component_change_BOXPLOT(base.batch, energy.batch, group.var = 'income.quint', var = S7_var)
  colnames(df2) <- c('run_id', 'Income_Quintile', S7_var)
  df <- merge(df, df2, by = c('run_id', 'Income_Quintile'))
}
rm(df2)

>>>>

```{r}
S7.var.list <- c('hh_income', 'equivalent_income',  # Income variables
                 'S7_housing_quality', 'S7_neighbourhood_safety', 'S7_physical_health', 'S7_mental_health', 'S7_labour_state', 'loneliness',  # SIPHER 7 variables
                 'ethnicity', 'age', 'region', 'job_sec', 'education_state', 'nkids_ind', 'housing_tenure', 'urban')
out.path.batch <- here::here('output', 'SIPHER7_batch/')

base.batch <- read_batch_out_all_years(out.path.batch, 'baseline', start.year = 2021, var.list = S7.var.list, verbose=FALSE)
int.batch <- read_batch_out_all_years(out.path.batch, 'energyDownlift', start.year = 2021, var.list = c('income_boosted', S7.var.list), verbose=FALSE)

int.name <- 'Living Wage'
```

@@

base.batch <- calc_age_group(base.batch)
int.batch <- calc_age_group(int.batch)
```

@@

base.batch <- calc_income_quintile(base.batch)
int.batch <- calc_income_quintile(int.batch)
```

@@

```{r}
income_equivinc_change_batch(base.batch, int.batch, int.name)
```

@@

```{r}
inc_change_batch_subpop(base.batch, int.batch, int.name, group.var = 'income.quint')
equivinc_change_batch_subpop(base.batch, int.batch, int.name, group.var = 'income.quint')
```

@@

  int.inc <- int.batch %>%
    select(pidp, time, run_id, hh_income, equivalent_income, all_of(group.var), weight) %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    summarise(int_income = weighted.mean(hh_income, w = weight, na.rm = TRUE),
              int_equiv_income = weighted.mean(equivalent_income, w = weight, na.rm = TRUE))

@@

  combined$diff_income <- combined$int_income - combined$base_income
  combined$diff_equiv_income <- combined$int_equiv_income - combined$base_equiv_income

@@

income_equivinc_percent_change_batch(base.batch, int.batch, int.name, group.var = 'income.quint')

@@

@@

  int.sub <- int.batch %>%
    select(pidp, time, run_id, equivalent_income, all_of(group.var), all_of(var)) %>%
    group_by(time, run_id, .data[[group.var]], .data[[var]]) %>%
    count() %>%
    group_by(time, run_id, .data[[group.var]]) %>%
    mutate(total = sum(n),
           int_prop = n / total) %>%
    select(-n, -total)

@@

```{r}
S7_pathway_change_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'S7_housing_quality')
S7_pathway_change_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'S7_neighbourhood_safety')
S7_pathway_change_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'S7_physical_health')
S7_pathway_change_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'S7_mental_health')
S7_pathway_change_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'loneliness')
```

## Cumulative Pathway Change

@@

```{r}
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'S7_housing_quality')
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'S7_neighbourhood_safety')
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'S7_physical_health')
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'S7_mental_health')
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'income.quint', var = 'loneliness')
```

@@

```{r}
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'age.group', var = 'S7_housing_quality')
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'age.group', var = 'S7_neighbourhood_safety')
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'age.group', var = 'S7_physical_health')
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'age.group', var = 'S7_mental_health')
S7_cumulative_pathwayChange_byGroup(base.batch, int.batch, group.var = 'age.group', var = 'loneliness')
```

@@

  if (var == 'S7_housing_quality') {
    base.batch$S7_housing_quality <- as.numeric(factor(base.batch$S7_housing_quality,
                                            levels = c('No to all',
                                                       'Yes to some',
                                                       'Yes to all'),
                                            labels = c(1, 2, 3)))
    int.batch$S7_housing_quality <- as.numeric(factor(int.batch$S7_housing_quality,
                                            levels = c('No to all',
                                                       'Yes to some',
                                                       'Yes to all'),
                                            labels = c(1, 2, 3)))
  } else if (var == 'S7_neighbourhood_safety') {
    base.batch$S7_neighbourhood_safety <- as.numeric(factor(base.batch$S7_neighbourhood_safety,
                                                 levels = c('Often',
                                                            'Some of the time',
                                                            'Hardly ever'),
                                                 labels = c(1, 2, 3)))
    int.batch$S7_neighbourhood_safety <- as.numeric(factor(int.batch$S7_neighbourhood_safety,
                                                 levels = c('Often',
                                                            'Some of the time',
                                                            'Hardly ever'),
                                                 labels = c(1, 2, 3)))
  } else if (var == 'loneliness') {
    base.batch$loneliness <- as.numeric(factor(base.batch$loneliness,
                                    levels = c('Often', 'Sometimes', 'Never'),
                                    labels = c(1, 2, 3)))
    int.batch$loneliness <- as.numeric(factor(int.batch$loneliness,
                                    levels = c('Often', 'Sometimes', 'Never'),
                                    labels = c(1, 2, 3)))
  }

@@

  p1 <- ggplot(data = merged2, aes(x = .data[[group.var]], y = sum.change)) +
    geom_boxplot(notch = FALSE) +
    geom_hline(yintercept = 0, linetype = 'dashed') +
    labs(title = paste0('Change in ', var, ' by ', group.var)) +
    ylab('% Change')

@@

```{r}
# simplified_component_change(base.batch, int.batch, group.var = 'income.quint', var = 'S7_housing_quality', print = TRUE)
# simplified_component_change(base.batch, int.batch, group.var = 'income.quint', var = 'S7_neighbourhood_safety', print = TRUE)
# simplified_component_change(base.batch, int.batch, group.var = 'income.quint', var = 'S7_mental_health', print = TRUE)
# simplified_component_change(base.batch, int.batch, group.var = 'income.quint', var = 'S7_physical_health', print = TRUE)
simplified_component_change(base.batch, int.batch, group.var = 'income.quint', var = 'loneliness', print = TRUE)
```

```{r}
# simplified_component_change_BOXPLOT(base.batch, int.batch, group.var = 'income.quint', var = 'S7_housing_quality', print = TRUE)
# simplified_component_change_BOXPLOT(base.batch, int.batch, group.var = 'income.quint', var = 'S7_neighbourhood_safety', print = TRUE)
# simplified_component_change_BOXPLOT(base.batch, int.batch, group.var = 'income.quint', var = 'S7_mental_health', print = TRUE)
# simplified_component_change_BOXPLOT(base.batch, int.batch, group.var = 'income.quint', var = 'S7_physical_health', print = TRUE)
simplified_component_change_BOXPLOT(base.batch, int.batch, group.var = 'income.quint', var = 'loneliness', print = TRUE)
```

@@

df <- simplified_component_change(base.batch, int.batch, group.var = 'income.quint', var = 'S7_housing_quality')
colnames(df) <- c('Income_Quintile', 'S7_housing_quality')
print("Completed for S7_housing_quality...")
for (S7_var in varlist) {
  df2 <- simplified_component_change(base.batch, int.batch, group.var = 'income.quint', var = S7_var)
  colnames(df2) <- c('Income_Quintile', S7_var)
  df <- merge(df, df2, by = c('Income_Quintile'))
  print(paste0("Completed for ", S7_var, '...'))
}
rm(df2)

@@

df <- simplified_component_change_BOXPLOT(base.batch, int.batch, group.var = 'income.quint', var = 'S7_housing_quality')
colnames(df) <- c('run_id', 'Income_Quintile', 'S7_housing_quality')
print("Completed for S7_housing_quality...")
for (S7_var in varlist) {
  df2 <- simplified_component_change_BOXPLOT(base.batch, int.batch, group.var = 'income.quint', var = S7_var)
  colnames(df2) <- c('run_id', 'Income_Quintile', S7_var)
  df <- merge(df, df2, by = c('run_id', 'Income_Quintile'))
  print(paste0("Completed for ", S7_var, '...'))
}
rm(df2)
